The Concurrent Calculi Formalisation Benchmark

### 2024-02-18 short-paper/short-paper.tex
coinduction, almost there both in Intro and in Design

\paragraph{Coinduction.} \todo{explain more what the challenge does.}
Coinduction is fundamental to establish properties of processes that
may never terminate, which is a typical situation.
Common uses for coinduction are to define strong and weak versions of
bisimulation, barbed notions of bisimulation, and prove trace equivalence. %,
%etc.
On the other hand, coinduction in proof assistants
is notoriously difficult: \lore{add here}
\cite{Castro-Perez2021} note that the most challenging
part of their mechanisation of multiparty session types
was the convoluted definition of infinite behaviours.

We base our challenge around bisimulation: a fundamental coinductively
established property for equating the
observed behaviour of potentially infinite processes. 
In this challenge, we study the notion of strong barbed bisimulation
(and bisimilarity, as in \cite{picalcbook})
over a fragment of the $\pi$-calculus and,
in particular, we focus on obtaining,
from strong barbed bisimilarity,
a congruence with respect to
parallel composition and substitution.

The introduction of bisimulation
is one of the first steps, when studying the
(observable) behaviour of process calculi, both
in classical textbooks (e.g., \cite{picalcbook})
and in almost all existent 
mechanisations, where we see a variety of
different formalisation choices.
As a significative example, we observe that,
while many authors rely
on the native support for coinduction offered by
proof assistants \cite{},
many others prefer a more set-theoretic
approach \cite{}.

@@

Our coinduction challenge concerns strong barbed bisimulation and
congruence.\alb{we motivate why strong vs weak, but not why barbed}
Though weak barbed congruence is a more common behavioural equivalence
w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
need to abstract over the number of internal transitions, thus
simplifying the theory.  We exclude delegation from the coinduction
challenge since it is orthogonal to our primary aim of exploring
coinductive proof techniques.

\lore{Need to integrate the following with the above.}
Different process-calculi
mechanisations
implement different techniques;
most of the time these 
are ad-hoc efforts to achieve a specific result.
\lore{cite and substantiate.}
In parallel, although most proof assistants
offer support for coinduction in their standard libraries,
research has mainly focused on the foundations
of the specific system, more than on
applicability.
\lore{cite and substantiate.}

>>>>

\paragraph{Coinduction.} \todo{explain more what the challenge does.}
\lore{I have tried this:
  the first paragraph motivates (``why is this difficult?''), while
  the second talks about the challenge.}
Coinduction is fundamental to establish properties of processes that
may never terminate, which is a typical situation.
Common uses for coinduction are to define strong and weak versions of
bisimulation, barbed notions of bisimulation, and prove trace equivalence. %,
%etc.
On the other hand, coinduction in proof assistants
is notoriously difficult and its representation
is non-uniform across development.
E.g., \cite{Castro-Perez2021} notes that the most challenging
part of their mechanisation of multiparty session types
was the convoluted definition of infinite behaviours;
\cite{Pohjola2022} points out that the main difficulty
of the proofs concerning bisimulation lies
in dealing with existential quantifiers; and
\cite{Bengtson2016} wonders whether parametrised
coinduction as developed in Coq \cite{Hur2013} would
have helped their Isabelle mechanisation.

We base our challenge around bisimulation: a fundamental coinductively
established property for equating the
observed behaviour of potentially infinite processes. 
In this challenge, we study the notion of strong barbed bisimulation
(and bisimilarity, as in \cite{picalcbook})
over a fragment of the $\pi$-calculus and,
in particular, we focus on obtaining,
from strong barbed bisimilarity,
a congruence with respect to
parallel composition and substitution.
%
%% The introduction of bisimulation
%% is one of the first steps, when studying the
%% (observable) behaviour of process calculi, both
%% in classical textbooks (e.g., \cite{picalcbook})
%% and in existing 
%% mechanisations
%% \cite{Thiemann2019,Honsell2001,Bengtson2016,Kahsai2008,Gay2020,
%% Hirschkoff1997,Bengtson2009,Maksimovic2015}.

%% , where we see a variety of
%% different formalisation choices.
%% As a significant example, we observe that,
%% while many authors rely
%% on the native support for coinduction offered by
%% proof assistants
%% \cite{Thiemann2019,Honsell2001,Bengtson2016,Kahsai2008,Gay2020},
%% many others prefer a more ``set-theoretic''
%% approach
%% \cite{Hirschkoff1997,Bengtson2009,Maksimovic2015}.

@@

\lore{Should we mention that Isabelle offers
  the command \texttt{coinduction\_upto}? Is it related?}
\todo{Insert discussion/related work focusing on existing techniques and how our benchmark design makes it possible to compare them.}
\lore{I have added something below... Any advice on what more we need?}

Our coinduction challenge concerns strong barbed bisimulation and
congruence.\alb{we motivate why strong vs weak, but not why barbed}
\lore{Is it not enough to quote the Sangiorgi/Walker book
and go by authority? :) }
Though weak barbed congruence is a more common behavioural equivalence
w.r.t.\ the \(\pi\)-calculus, here we prefer strong equivalences to avoid the
need to abstract over the number of internal transitions, thus
simplifying the theory.  We exclude delegation from the coinduction
challenge since it is orthogonal to our primary aim of exploring
coinductive proof techniques.

The introduction of strong barbed bisimulation
is one of the first steps, when studying the
(observable) behaviour of process calculi, both
in classical textbooks (e.g., \cite{picalcbook})
and in existing 
mechanisations, where we see a variety of
different formalisation choices,
sometimes ad-hoc for the specific
result or development.
As a significant example, we observe that,
while many authors rely
on the native support for coinduction offered by
proof assistants
\cite{Honsell2001,Bengtson2016,Kahsai2008,Thiemann2019,Gay2020},
many others prefer a more ``set-theoretic''
approach
\cite{Hirschkoff1997,Bengtson2009,Maksimovic2015,Pohjola2022}.

%% \lore{Need to integrate the following with the above.}
%% Different process-calculi
%% mechanisations
%% implement different techniques;
%% most of the time these 
%% are ad-hoc efforts to achieve a specific result.
%% \lore{cite and substantiate.}
%% In parallel, although most proof assistants
%% offer support for coinduction in their standard libraries,
%% research has mainly focused on the foundations
%% of the specific system, more than on
%% applicability.
%% \lore{cite and substantiate.}
